---
title: "R - GPU Programming for All with gpuR"
output:
  html_document:
    df_print: paged
---

Source: http://www.parallelr.com/r-gpu-programming-for-all-with-gpur/


```{r results='hold', warning=FALSE}
library(gpuR)

ORDER = 1024
 
A = matrix(rnorm(ORDER^2), nrow=ORDER)
B = matrix(rnorm(ORDER^2), nrow=ORDER)
gpuA = gpuMatrix(A, type="double")
gpuB = gpuMatrix(B, type="double")
 
C = A %*% B
gpuC = gpuA %*% gpuB
 
all.equal(C, gpuC[])
# TRUE
```

The gpuMatrix object points to a matrix in RAM which is then computed by the GPU when a desired function is called.  This avoids R’s habit of copying the memory of objects.  For example:

```{r results='hold'}
library(pryr)
 
# Initially points to same object
x = matrix(rnorm(16), 4)
y = x
 
address(x)
# "0x77ebdd0"

address(y)
# "0x77ebdd0"
 
# But once modify the second object it creates a copy
y[1,1] = 0
 
address(x)
# "0x77ebdd0"

address(y)
# "0x75aef00"
```

In contrast, the same syntax for a gpuMatrix will modify the original object in-place without any copy.

```{r results='hold'}
library(pryr)
library(gpuR)
 
# Initially points to same object
x = gpuMatrix(rnorm(16), 4, 4)
y = x
 
x@address
# <pointer: 0x000000001df4d550>
 
y@address
# <pointer: 0x000000001df4d550>
 
# Modification affects both objects without copy
y[1,1] = 0
 
x@address
# <pointer: 0x000000001df4d550>
 
y@address
# <pointer: 0x000000001df4d550>
```

Each new variable assigned to this object will only copy the pointer thereby making the program more memory efficient.  However, the `gpuMatrix` class does still require allocating GPU memory and copying data to device for each function call. The most commonly used methods have been overloaded such as  %*%, +, -, *, /, crossprod, tcrossprod, and trig functions among others.  In this way, an R user can create these objects and leverage GPU resources without the need to know a bunch more functions that would break existing algorithms.

